Inferring the effective start dates of non-pharmaceutical interventions during COVID-19 outbreaks

### 2020-05-20 ms/ms.tex
added WAIC

@@

In contrast, in most investigated countries (e.g., Spain and France), we infer reduced transmission rates even before official lockdowns were implemented (\autoref{fig:fig3}).
This early response might be due to adoption of social distancing and similar behavioral adaptations in parts of the population, maybe in response to increased risk perception due to domestic or international COVID\=/19-related reports.
This finding may also suggest that severe NPIs, such as lockdowns, were unnecessary, and that less extreme measures adopted by the population could have been sufficient for epidemic control. These less extreme measures may have been implemented due to government recommendations, media coverage, and social networks, rather than official NPIs.
\hl{check if this is true} Indeed, the evidence supports a change in transmission dynamics (i.e. a model with $\tau$) even for Sweden, in which a lockdown was not implemented, suggesting that lockdowns may not be necessary if other NPIs are adopted early enough during the outbreak~\citep{Banholzer2020} (Sweden banned public events on Mar~12, encouraged social distancing on Mar~16, and closed schools on Mar~18~\citep{Flaxman2020}.)

@@

@@

    \textbf{Figure S1. Posterior prediction check plots}

@@

%%% Fig Switzerland %%%
\begin{figure}[h]
    \centering
        \includegraphics[width=0.75\textwidth]{../figures/posterior/Switzerland_τ_posterior.pdf}
    \caption{
	\textbf{Late effect of non-pharmaceutical interventions in Switzerland.}
    Posterior distribution of $\tau$, the effective start date of NPI, is shown as a histogram of MCMC samples. Red line shows the official last NPI date $\tau^*$. Black line shows the estimated $\hat{\tau}$. Shaded area shows a 95\% credible interval (area in which $P(|\tau - \hat{\tau}| \mid \vec{X}) = 0.95$). 
    }
    \label{fig:Switzerland}
\end{figure}

>>>>

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

In contrast, in most investigated countries (e.g., Spain and France), we infer reduced transmission rates even before official lockdowns were implemented (\autoref{fig:fig3}).
An early effective date might be due to adoption of social distancing and similar behavioral adaptations in parts of the population, maybe in response to increased risk perception due to domestic or international COVID\=/19-related reports.
This finding may also suggest that severe NPIs, such as lockdowns, were unnecessary, and that less extreme measures adopted by the population could have been sufficient for epidemic control.
These less extreme measures may have been implemented due to government recommendations, media coverage, and social networks, rather than official NPIs.
Indeed, the evidence supports a change in transmission dynamics (i.e. a model with free $\tau$) even for Sweden~(\autoref{fig:ppc}), in which a lockdown was not implemented\footnote{Sweden banned public events on Mar~12, encouraged social distancing on Mar~16, and closed schools on Mar~18~\citep{Flaxman2020}.}, suggesting that lockdowns may not be necessary if other NPIs are adopted early enough during the outbreak~\citep{Banholzer2020}.

@@

	\subcaption{Wuhan, China}    
        \includegraphics[width=\textwidth]{../figures/ppc/fixed/Wuhan_ppc.pdf}
        \includegraphics[width=\textwidth]{../figures/ppc/free/Wuhan_ppc.pdf}

@@

    \textbf{Figure S2. Posterior prediction check plots}

@@

% Table WAIC %%%
\begin{table}[h]
\centering
\pgfplotstabletypeset[
    col sep=comma,
    string type,
    every head row/.style={before row=\hline,after row=\hline},
    every last row/.style={after row=\hline},
    ]{../figures/Table-WAIC.csv}
\caption{
\textbf{WAIC values for the different models.}
	WAIC (widely applicable information criterion)~\citep{gelman2013bayesian} values for models  with: \emph{No}: no $\tau$ at all; \emph{Fixed}: $\tau$ fixed at the official last NPI date $\tau^*$; and \emph{Free}: free $\tau$. WAIC values are scaled as a deviance measure: lower values imply higher predictive accuracy.
}
\label{table:table1}
\end{table}
